---
output: 
  pdf_document:
  #stevetemplates::article2:
  #  citation_package: natbib
    dev: cairo_pdf
    keep_tex: true
biblio-style: apsr
title: "Title of the semester project"
graphics: yes
author:
- name: Thuy-My Ngo 
  affiliation: University of Geneva
- name: Doriane Pfister
  affiliation: University of Geneva
- name: Léonie Lefebvre
  affiliation: University of Geneva
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
#mainfont: cochineal
#sansfont: Linux Biolinum O
fontsize: 11pt
# doublespacing: TRUE
endnote: no
# pandocparas: TRUE
sansitup: FALSE
bibliography: biblio.bib 
header-includes:
  - \usepackage{longtable}
  - \usepackage{hologo}
  - \LTcapwidth=.95\textwidth
  - \linespread{1.05}
  - \usepackage{hyperref}
---
```{r functions, echo=FALSE}
# Generate Bivariate Data with Nonlinear Relationship
gen_nonlinear <- function(n, muX, muY, sX, sY, angle){
  rotmat <- matrix(c(cos(angle), -sin(angle), sin(angle), cos(angle)),2,2)
  x <- rnorm(n, muX, sX)
  y <- x**2 + rnorm(n, muY, sY)
  df <- cbind(x,y)
  df <- df %*% rotmat
  df <- as.data.frame(df)
  names(df) <- c("x", "y")
  return(df)
}

# True correlation value for the nonlinear setting
cor_nonlinear <- function(muX, muY, sX, sY, angle){
  # muX and muY have no effect
  sQ <- sqrt(cos(angle)**2 * (2*sX**4 + 4*sX**2*muX**2 +  sY**2) + sin(angle)**2 * sX**2 + 2 * cos(angle)*sin(angle)*(2*sX**2*muX))
  sR <- sqrt(sin(angle)**2 * (2*sX**4 + 4*sX**2*muX**2 + sY**2) + cos(angle)**2 * sX**2 - 2 * cos(angle)*sin(angle)*(2*sX**2*muX))
  cov <- sin(angle)*cos(angle) * (sX**2 - (2*sX**4 + 4*sX**2*muX**2 + sY**2)) +
    (cos(angle)**2 - sin(angle)**2) * (2*sX**2*muX)
  cor <- round(cov/(sQ*sR), 4)
  return(cor)
}

```

# `RMarkdown` basics

This is a citations: @efron1992bootstrap.   


This is a displayed but not evaluated `R` code chunk

This is an `R` code chunk, not displayed but evaluated.

```{r, fig.width=5, fig.height=4, fig.align='center', echo =F}
data=rnorm(100)
par(mfrow=c(2,1))
par(mar=c(2,4,1,2))
plot(density(data), main = "", ylab ="", xlab ="")
boxplot(data, horizontal = T)
par(mfrow=c(1,1))
mean_data = mean(data)
```

This is an inline `R` code: Hence, the mean of the data is of `r mean_data`.

This is a \hologo{LaTeX} equation 

$$
f(x)=\frac{1}{\sigma \sqrt{2 \pi}} \mathrm{e}^{-\frac{1}{2}\left(\frac{x-\mu}{\sigma}\right)^{2}}
$$

This is a inline \hologo{LaTeX} equation: $\frac{1}{n} \sum_{i=1}^{n} a_{i}=\frac{a_{1}+a_{2}+\cdots+a_{n}}{n}$

# Introduction

La moyenne vaut `r round(mean_data,2)`

# Analysis

# Description of the task

# Motivation

# Results: description and interpretation
```{r nonlinear_q1, echo=FALSE, fig.align = 'center',fig.cap = '**Eyebrow 1** This is the population correlation with angle = 0'}

#Ex2 question 1 (a)
#How does the angle parameter visually affect the data?
#angle = 0 

# Parameters:
SAMPLESEED <- 9886
SIMULATIONSEED <- 1021
n <- 500
muX <- 0
muY <- -1
sX  <- 1.8
sY  <- 0.5
rho <- 0.84
out <- 0.03
dev <- 3
angle <- 0

set.seed(SAMPLESEED)
df <- gen_nonlinear(n=n, muX=muX, muY=muY, sX=sX, sY=sY, angle=pi)
# population correlation in this setting 
rho <- cor_nonlinear(muX=muX, muY=muY, sX=sX, sY=sY, angle=angle)

plot(df, main=paste("Population nonlinear correlation = ", round(rho, 3)))

```


```{r nonlinear_q1, echo=FALSE, fig.align = 'center',fig.cap = '**Eyebrow 2** This is the population correlation with angle = -0.45'}

#Ex2 question 1 (a)
#How does the angle parameter visually affect the data?
#angle = -0.45 

# Parameters:
SAMPLESEED <- 9886
SIMULATIONSEED <- 1021
n <- 500
muX <- 0
muY <- -1
sX  <- 1.8
sY  <- 0.5
rho <- 0.84
out <- 0.03
dev <- 3
angle <- -0.45

set.seed(SAMPLESEED)
df <- gen_nonlinear(n=n, muX=muX, muY=muY, sX=sX, sY=sY, angle=pi)
# population correlation in this setting 
rho <- cor_nonlinear(muX=muX, muY=muY, sX=sX, sY=sY, angle=angle)

plot(df, main=paste("Population nonlinear correlation = ", round(rho, 3)))

```


# Were these results expected: discussion

# Statistical methods used

# Acquired skills during the term project

# Additional element

# Conclusion